Advection Versus Diffusion in Intraventricular Transport

### 2025-02-18 paper/main.tex
last version before template change
Wiley template

\documentclass[fleqn]{wlscirep}
\usepackage[utf8]{inputenc}
\usepackage{lipsum}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{siunitx}
\usepackage{physics}
\usepackage{stmaryrd} % double brackets for jump operator
\usepackage{nameref}
\usepackage{booktabs}
\usepackage{bm}
\usepackage{amsopn}
\usepackage[skip=10pt]{caption}
\usepackage{color}
\usepackage{xcolor,colortbl}
\usepackage[nameinlink, capitalise]{cleveref}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{tablefootnote} % for table footnotes
\usepackage{float} % figure placement

% Bibliography tools
\usepackage{csquotes}

% Set some cref options
\newcommand{\crefrangeconjunction}{--} % Change crefrange from "to" to dashed

%% Some new commands
% Math
\newcommand{\cbar}{\overline{c}}

% Partial derivatives
\newcommand{\pdifft}[1]{\frac{\partial  #1}{\partial t}}
\newcommand{\pdiffx}[1]{\frac{\partial #1}{\partial x}}
\newcommand{\pdiffy}[1]{\frac{\partial #1}{\partial y}}
\newcommand{\pdiffxi}[1]{\frac{\partial #1}{\partial x_i}}
\newcommand{\pdiffxj}[1]{\frac{\partial #1}{\partial x_j}}
\newcommand{\secpdiffx}[1]{\frac{\partial^2 #1}{\partial x^2}}
\newcommand{\secpdiffxi}[1]{\frac{\partial^2 #1}{\partial x_i \partial x_i}}
\newcommand{\secpdiffxj}[1]{\frac{\partial^2 #1}{\partial x_j \partial x_j}}

% Norms
\newcommand{\normltwo}[1]{{ \vert\vert#1\vert\vert}_{L^2}}
\newcommand{\normltwovec}[1]{{ \vert\vert#1\vert\vert}_{\mathbf{L}^2}}
\newcommand{\normlinf}[1]{{\vert\vert#1\vert\vert}_{L^{\infty}}}
\newcommand{\Hnorm}[1]{\norm{#1}_{\mathbf{H}^1}}
\newcommand{\Hdivnorm}[1]{\norm{#1}_{\mathbf{H}(\mathrm{div};\Omega)}}

% Integrals
\newcommand{\dx}{\, \mathrm d\bm{x}}
\newcommand{\intO}[1]{\int_{\Omega}#1 \, \mathrm d\bm{x}}
\newcommand{\intG}[1]{\int_{\Gamma}#1 \, \mathrm ds}
\newcommand{\intF}[1]{\int_{\mathcal{F}_k}#1 \, \mathrm dS}
\newcommand{\intGnf}[1]{\int_{\Gamma_{\mathrm{nf}}}#1 \, \mathrm ds}
\newcommand{\intGinj}[1]{\int_{\Gamma_{\mathrm{inj}}}#1 \, \mathrm ds}
\newcommand{\intGin}[1]{\int_{\Gamma_{\mathrm{in}}}#1 \, \mathrm ds}
\newcommand{\intGout}[1]{\int_{\Gamma_{\mathrm{out}}}#1 \, \mathrm ds}
\newcommand{\intGinjout}[1]{\int_{\Gamma_{\mathrm{inj}}\cup\Gamma_{\mathrm{out}}}#1 \, \mathrm ds}
\newcommand{\intGs}[1]{\int_{\Gamma_s}#1 \, \mathrm ds}
\newcommand{\intGc}[1]{\int_{\Gamma_c}#1 \, \mathrm ds}
\newcommand{\intGp}[1]{\int_{\Gamma_p}#1 \, \mathrm ds}

% Domain and boundaries
\newcommand{\Gs}{\Gamma_{s}}
\newcommand{\Gc}{\Gamma_{c}}
\newcommand{\Gp}{\Gamma_{p}}
\newcommand{\Ginj}{\Gamma_{\mathrm{inj}}}
\newcommand{\Gnf}{\Gamma_{\mathrm{nf}}}
\newcommand{\Gin}{\Gamma_{\mathrm{in}}}
\newcommand{\Gout}{\Gamma_{\mathrm{out}}}

% Discontinuity operators
\newcommand{\avg}[1]{\{#1\}}
\newcommand{\jump}[1]{\llbracket#1\rrbracket}

% Bold faced math characters
\newcommand{\ff}{\mathbf{f}}
\newcommand{\nn}{\mathbf{n}}
\newcommand{\rr}{\mathbf{r}}
\newcommand{\uu}{\mathbf{u}}
\newcommand{\vv}{\mathbf{v}}
\newcommand{\ww}{\mathbf{w}}
\newcommand{\xx}{\bm{x}}
\newcommand{\JJ}{\mathbf{J}}
\newcommand{\VV}{\mathbf{V}}
\newcommand{\bsig}{\bm{\sigma}}
\newcommand{\bsigpar}{\hat{\bsig}_{\parallel}}
\newcommand{\bsigperp}{\hat{\bsig}_{\perp}}
\newcommand{\beps}{\bm{\varepsilon}}
\newcommand{\btau}{\bm{\tau}}

% Comments
\newcommand{\lyng}[1]{\textcolor{blue}{#1}}
\newcommand{\mer}[1]{\textcolor{magenta}{#1}}
\newcommand{\mk}[1]{\textcolor{olive}{MK: #1}}

\newcommand{\fixme}[1]{\textcolor{red}{#1}}

% Title
\title{The role of advection versus diffusion in intraventricular transport}
\author[1,*]{Halvor Herlyng}
\author[1]{Ada J.~Ellingsrud}
\author[1]{Miroslav Kuchta} 
\author[2]{Inyoung Jeong} % OK?
\author[1,3,+]{Marie E.~Rognes} % OK order?
\author[2,+,*]{Nathalie Jurisch-Yaksi}
\affil[1]{Department of Numerical Analysis and Scientific Computing,
Simula Research Laboratory, Oslo, Norway}
\affil[2]{Department of Clinical and Molecular Medicine,
Norwegian University of Science and Technology, Trondheim, Norway}
\affil[3]{K.~G.~Jebsen Center for Brain Fluid Research}
\affil[*]{\href{mailto:hherlyng@simula.no}{hherlyng@simula.no},
\href{mailto:nathalie.jurisch-yaksi@ntnu.no}{nathalie.jurisch-yaksi@ntnu.no}}
\affil[+]{these authors contributed equally to this work}
% Optional PDF information
\ifpdf
\hypersetup{
  pdftitle={zebrafish-cilia-csf},
  pdfauthor={H.~Herlyng~\emph{et al.}}
}
\fi

\begin{abstract}
    This paper presents ...
    %Flow of cerebrospinal fluid in zebrafish brain ventricles is studied in this work. Motivated by experimental observations, we consider motile cilia and cardiac beating as the driving flow mechanisms.
    %A mathematical model of the flow is developed and coupled with a transport model to study the transport of molecules inside the brain ventricles. 
    %The Stokes equations are used to model the flow, with the motile cilia acting as a force on the fluid through a stress boundary condition that allows slip at the wall. 
    %Transport is modeled with an advection-diffusion equation coupled to the velocity field obtained from solving the Stokes equations.
    %The flow and transport models are discretized with finite element methods. For the Stokes equations we employ a non-conforming discretization based on elements of the Brezzi-Douglas-Marini family. 
    %The advection-diffusion equation is discretized with discontinuous Galerkin elements. 
    %We implement and solve the discretized equations numerically to simulate flow and transport in a zebrafish brain ventricles geometry obtained with medical imaging.
\end{abstract}
\begin{document}
\maketitle
% Introduction:
% 1. What are cilia and why are they important?
% 2. Previous studies of cilia flow
% 3. What problem do we want to solve
% 4. The model and work done in this paper
\section*{Introduction}
Brain development and function depend on numerous regulatory processes,
including interactions with cerebrospinal fluid
(CSF). CSF is a clear liquid produced by the choroid plexus
that flows through the ventricular system and subarachnoid space.
By delivering growth factors, nutrients, neurotransmitters and metabolites
and removing waste products, CSF regulates brain development and homeostasis.
Disturbances of CSF circulation can lead to brain disorders, such as
hydrocephalus where accumulation of CSF often requires emergency surgical treatment.

@@

\section*{Methods}
\subsection*{Image-based computational geometries of zebrafish brain ventricles}
To model flow and transport in zebrafish larval brain ventricles, we generated a
3D representation of the brain ventricular walls using confocal imaging data of a zebrafish embryo
injected intraventricularly with a 70 kDa dye at 2 dpf~\cite{Olstad2019CiliaryDevelopment}. 
From the imaging data, we constructed a computational mesh of the ventricles (\Cref{fig:fig1}d).
The interior volume was meshed with fTetWild~\cite{Hu2020FastWild}. 
The \emph{standard} mesh consisted of 132 134 tetrahedral cells and had a
maximal (minimal) edge length of 14.9 \textmu m (4.68 \textmu m). 
At 2 dpf, the ventricular system consists mainly of 3 cavities: the telencephalic (anterior),
the di-/mesencephalic (middle) and the rhombencephalic (posterior) ventricles, which are connected by ducts. 
Parts of the ventricular surfaces were marked to be lined with
motile cilia~\cite{Olstad2019CiliaryDevelopment} (\Cref{fig:fig1}d, magenta markers).
\begin{figure}%[hbt!]
    \centering
    \includegraphics[width=\linewidth]{graphics/figure1.png}
    \caption{
    \textbf{a.} Schematic illustration of the two cerebrospinal fluid (CSF) flow components modeled,
    motile cilia and cardiac pulsations, and their contributions to the flow patterns.
    \textbf{b.} Particle tracking data visualizing flow directionality in the middle ventricle.
    \textbf{c.} Flow velocity magnitudes in the middle ventricle from particle tracking data.
    \textbf{d.} Computational mesh with marked facets indicating regions of motile cilia (magenta) and
    anterior/posterior facets (blue).
    \textbf{e.} Streamlines of the (steady-state) CSF flow field simulated with the
    cilia-only/no-cardiac flow model.
    \textbf{f.} Streamlines of the CSF flow field simulated with the cardiac-only/no-cilia flow
    model at a time 25\% into the cardiac cycle. %t=0.112613 s, period=0.45045 s
    \textbf{g.} Streamlines simulated with the baseline flow model at
    a time 25\% into the cardiac cycle. %t=0.112613 s
    \textbf{h.} Streamlines simulated with the baseline flow model at
    a time 50\% into the cardiac cycle. %t=0.225225 s
    \textbf{i.} Streamlines simulated with the baseline flow model at
    a time 75\% into the cardiac cycle. %t=0.315315 s
    \textbf{j.} Illustration of the cardiac-related normal pressure forcing imposed on the
    anterior/posterior boundaries. Time instants of panels \textbf{f, h--j} are indicated.
    }
    \label{fig:fig1}
\end{figure}

\subsection*{Computational CSF dynamics}
The beating motion of the cilia generates CSF flow with steady rotational
structures~\cite{Olstad2019CiliaryDevelopment}
within the brain ventricles (\Cref{fig:fig1}a--c).
In addition, cardiac pulsations induce pulsatile CSF flow.
We model this flow of CSF in the brain ventricles by the time-dependent,
incompressible Stokes equations which read as follows.
Find the CSF velocity $\uu = \uu(\xx, t) = (u_x(\xx, t), u_y(\xx, t), u_z(\xx, t))$
and the CSF pressure $p = p(\xx, t)$ for $\xx \in \Omega$ and time $t>0$, such that
\begin{subequations}
    \begin{align}
      \rho \pdifft{\uu} - \nabla \cdot \bsig &= \mathbf{0}
      \quad \text{in } \Omega,\\
      \nabla \cdot \uu &= 0
      \quad  \text{in } \Omega,
  \end{align}
  \label{eq:stokes_eqs}% 
\end{subequations}%
where the stress tensor is defined as $\bsig(\uu, p) = 2\mu\beps(\uu) - p\mathbf{I}$
and $\beps(\uu) = \frac{1}{2}\left(\nabla \uu + (\nabla\uu)^T\right)$ is the strain-rate tensor. 
Bold-face characters denote vectors or tensors, and $\mathbf{I}$ is the identity tensor in three dimensions.
Furthermore, $\rho=1000 \ \mathrm{kg/m^3}$ is the CSF density
and $\mu=0.70 \ \mathrm{mPa\cdot s}$ is the dynamic viscosity
of the CSF~\cite{Bloomfield1998EffectsFluid}.
We introduce the tangential traction $\bsigpar$ as the
tangential component of the traction $\hat{\bsig}=\bsig\nn$:
\begin{equation*}
    \bsigpar = P_{\nn}(\hat{\bsig}) = (\mathbf{I} - \nn\otimes\nn)\hat{\bsig},
\end{equation*}
where $\nn$ is the outer unit normal of the boundary surface $\Gamma = \partial\Omega$,
and $P_{\nn}(\rr)$ is the tangential projection of a vector $\rr$ onto $\Gamma$.
We consider three types of boundary conditions:
(i) \emph{tangential traction} (no flow normal to the boundary
$\uu \cdot \nn = 0$ with a force $\btau$ applied tangentially on the boundary $\bsigpar = \btau$),
(ii) \emph{normal pressure} (a pressure $\Tilde{p}$ applied in
the normal direction: $(\mu\nabla\uu - p\mathbf{I})\nn = \Tilde{p}(t)\nn$),
and (iii) \emph{free slip} (no flow normal to the boundary
$\uu \cdot \nn = 0$ and no tangential forces $\bsigpar = {\bf 0}$).
Unless noted otherwise, the free-slip condition is applied on the
boundary. We assume that the system starts at rest: $\uu(\xx, t=0) = \mathbf{0}$.

\subsection*{Cilia-driven flow}
On the ciliated regions of the ventricle walls $\Gc$ (\Cref{fig:fig1}d, magenta markers),
we impose a constant-in-time tangential traction $\bsigpar=\btau$ to represent the net forces
of the cilia acting on the CSF.
We set $\btau(\xx, \rr) = \tau \lambda(\xx) P_{\nn}(\rr)$, where the function $\lambda(\xx)$
and the sign of the vector $\rr=\pm(1, 0, 1)$ varies depending on the
cilia region (see~\Cref{subsec:appendixA1}).
Both $\rr$ and the value $\tau=0.65$ mPa were chosen by 
model calibration using experimental data~\cite{Olstad2019CiliaryDevelopment}.
The data indicate speeds proximate to the ventricular walls of $27.4 \pm 5.4$ \textmu m/s and
$2.6 \pm 0.6$ \textmu m/s in the dorsal and ventral regions of the middle ventricle, respectively.
In the anterior ventricle, the speeds were estimated at $4.7 \pm 1.3$ \textmu m/s.
We calibrated $\tau$ heuristically by comparing these experimental velocity values
with the simulated, maximum velocity magnitude.

\subsection*{Flow induced by cardiac pulsations}\label{subsec:pressure_bcs}
To drive a cardiac-induced pulsatile flow inside the brain ventricles, we set $\Tilde{p}(t)=0$ and
$\Tilde{p}(t)=-A\sin{\omega t}$ on anterior and posterior parts, respectively, of the
boundary $\Gp$~(\Cref{fig:fig1}d, blue markers).
Based on a measured cardiac frequency of $f=2.22$ Hz~\cite{Olstad2019CiliaryDevelopment},
the angular cardiac frequency $\omega=2\pi f=6.97$ rad/s.
The amplitude $A=1.5$ mPa determines the magnitude of the pulsatile flow,
and was calibrated based on experimental data~\cite{Olstad2019CiliaryDevelopment}.
Similar to $\tau$, the amplitude $A$ was chosen by comparing experimental and simulated velocities.
Particle tracking data of pulsatile motion was used to determine the contributions of the
cardiac pulsations to the velocity magnitude,
as compared to the effect of cilia motion on the flow (see~\Cref{sec:appC_particle_transport}).

\subsection*{Computational model of solute transport after photoconversion}
We simulate transport of photoconverted proteins within the brain ventricles by
modeling transport of a solute with concentration $c(\xx, t)$ for $\xx\in\Omega$ and
time $t > 0$ by the advection-diffusion equation
\begin{equation}
    \pdifft{c} + \nabla\cdot\JJ = 0 \quad \text{in } \Omega,
    \label{eq:adv_diff_strong}
\end{equation}
where the total concentration flux $\mathbf{J}$ is assumed to consist of an advective and a diffusive flux:
\begin{equation*}
    \JJ = c\uu - D\nabla c .
\end{equation*}
Here, $\uu$ is the CSF velocity field, governed by the Stokes equations~\eqref{eq:stokes_eqs},
and $D$ is the molecular diffusion coefficient of the solute.
Initially, we set $c(\xx, 0)=0$ for all $\xx\in\Omega$.
The ventricular walls are assumed to be impermeable (admitting \emph{no flux}: $\JJ\cdot\nn=0$),
except at the anterior/posterior boundary $\Gamma_p$.
At the anterior/posterior boundary, we permit the solute concentration to be advected into
and out of the domain in a manner that ensures conservation of mass
(see~\Cref{subsec:adv_diff_eq_weak_form} for details).

\subsection*{Estimation of diffusion coefficients via the Stokes-Einstein relation}
Motivated by an interest in molecules that are significant in neural development,
we study the transport of molecules resembling the Dendra2 Fluorescent Protein (Dendra2)
used in this study, the Starmaker+Green Fluorescent Protein (STM-GFP) reported in
Jeong~\emph{et al.}~\cite{Jeong2024TheZebrafish}, and exosomes.
To estimate the diffusion coefficients of these, when not
available in the literature, we use the Stokes--Einstein relation:
\begin{equation}
    D = \frac{k_B T}{6\pi \mu R},
    \label{eq:D_stokes_einstein}
\end{equation}
which is an expression for the diffusion coefficient of a spherical
molecule suspended in a viscous fluid~\cite{Einstein1905UberTeilchen}.
In~\eqref{eq:D_stokes_einstein}, $k_B = \num{1.38e-23}$ J/K is the Boltzmann constant,
$T = 310$ K is the absolute temperature, $\mu$ is the dynamic viscosity of the
solvent (in this case the CSF), and $R$ is the radius of the molecule.
For the exosomes, we calculate $D$ using~\eqref{eq:D_stokes_einstein}
with $R=200$ nm. This radius covers the slowest diffusing exosomes,
since the smaller exosomes would have diffusion coefficients similar to the other molecules.
For STM-GFP, we estimated the diffusion coefficient by extrapolating
the experimentally observed diffusion coefficient of
GFP~\cite{Swaminathan1997PhotobleachingDiffusion, Potma2001ReducedCells},
assuming that if $D$ scales with $1/R$ according to~\eqref{eq:D_stokes_einstein},
$D$ would scale with the cube root of the molecule mass~\cite{Goodhill1997DiffusionGuidance}.
The resulting diffusion coefficients used in the simulations
are $D_1$ (exosomes), $D_2$ (STM-GFP) and $D_3$ (Dendra2)~(\Cref{tab:diff_coeff_table}).

\begin{table}[h!]
    \centering
    \begin{tabular}{l|ccc}
        \toprule
        Molecule & Mass [kDa] & $D \ [\mathrm{m^2/s}]$ & Reference \\ 
        \midrule 
        Dendra2 Fluorescent Protein & 25.6~\cite{Gurskaya2006EngineeringLight}
        & $\num{1.15e-10}$  & \cite{GuraSadovsky2017MeasurementExpansion}\\
        GFP (saline aqueous solution) & 26.9~\cite{UniProt2024GFPUniProtKB}
        & $\num{8.70e-11}$ & \cite{Swaminathan1997PhotobleachingDiffusion, Potma2001ReducedCells} \\
        STM-GFP & 66.2~\cite{UniProt2024StmUniProtKB} + 26.9~\cite{UniProt2024GFPUniProtKB}
        & $\num{5.75e-11}$ & * \\
        Exosomes (radius of 200 nm) & -- & $\num{1.63e-12}$ & \textdagger \\
        \bottomrule
    \end{tabular}
    \caption{Molecular mass and diffusion coefficients of simulated solutes.
    STM and GFP denote Starmaker Protein and Green Fluorescent Protein, respectively.
    (*) The value was extrapolated based on the value of $D$ for
    GFP reported in the literature~\cite{Swaminathan1997PhotobleachingDiffusion, Potma2001ReducedCells},
    assuming that $D$ scales with the cube root of the molecule mass.
    (\textdagger) The value was estimated using the Stokes-Einstein relation~\eqref{eq:D_stokes_einstein}.}
    \label{tab:diff_coeff_table}
\end{table}

>>>>

\documentclass{WileyMSP-template}

\begin{document}

\pagestyle{fancy}
\rhead{\includegraphics[width=2.5cm]{vch-logo.png}}

\title{Advection Versus Diffusion in Intraventricular Transport}

\maketitle

% Author: Please give full first and last names for authors and include * after the name of all corresponding authors

\author{Halvor Herlyng*}
\author{Ada J.~Ellingsrud}
\author{Miroslav Kuchta}
\author{Inyoung Jeong}
\author{Marie E.~Rognes} \\
\author{Nathalie Jurisch-Yaksi*}

% Affiliations: Please provide adacemic titles (Prof. or Dr.) for all authors where applicable, and include an institutional email address for all corresponding authors
\begin{affiliations}
H.~Herlyng, Dr.~A.~J.~Ellingsrud, Dr.~M.~Kuchta, Dr.~M.~E.~Rognes\\
Department of Numerical Analysis and Scientific Computing,
Simula Research Laboratory, Kristian Augusts gate 23, 0164 Oslo, Norway\\
Email Address: hherlyng@simula.no

Dr.~M.~E.~Rognes \\
K.~G.~Jebsen Center for Brain Fluid Research, Oslo, Norway
\\
Dr.~I.~Jeong, Dr.~N.~Jurisch-Yaksi\\
Department of Clinical and Molecular Medicine,
Norwegian University of Science and Technology, Erling Skjalgsons gate 1, 7491 Trondheim, Norway \\
Email Address: nathalie.jurisch-yaksi@ntnu.no

\end{affiliations}

% Keywords: Please provide a minimum of three and a maximum of seven keywords, separated by commas

\keywords{cerebrospinal fluid, brain ventricles, fluid flow, transport}

% Abstract should be written in the present tense and impersonal style (i.e., avoid we), and be at most 200 words long
\begin{abstract}

Please insert your abstract here

\end{abstract}

% Text: Please use section headings and subheadings as specified below. For communications, all section headings apart from Experimental Section should be removed
% Please make the first reference to a display item bold: \textbf{Figure 1}
% Do not abbreviate Figure, Equation, etc.; display items are always singular, i.e., Figure 1 and 2.
% Equations are always singular, i.e., Equation 1 and 2, and should be inserted using the {equation} environment, not as graphics
% Please do not use footnotes in the text, additional information can be added to the Reference list.

%\section{First Section}

%\subsection{First Subsection}

%\subsubsection{First Sub Subsection}

%\threesubsection{First lowest-level subsection}

\section{Introduction}
Brain development and function depend on numerous regulatory processes,
including interactions with cerebrospinal fluid
(CSF). CSF is a clear liquid produced by the choroid plexus
that flows through the ventricular system and subarachnoid space.
By delivering growth factors, nutrients, neurotransmitters and metabolites
and removing waste products, CSF regulates brain development and homeostasis.
Disturbances of CSF circulation can lead to brain disorders, such as
hydrocephalus where accumulation of CSF often requires emergency surgical treatment.

@@

%%--------------RESULTS--------------%%
\section{Results}
\subsection{Motile cilia induce flow compartmentalization within the brain ventricles}

We model the compartmentalized and pulsatile CSF flow observed in vivo~(\Crefpanels{fig:fig1}{b, c})
by the Stokes equations, and solve the equations numerically using a finite element method.
To characterize CSF flow patterns predicted by
computational fluid dynamics within the brain ventricles, we first
consider the simulated CSF velocity and pressure when only modeling
cilia motion, disregarding cardiac pressure pulsations. The
bidirectional traction induced by the motile cilia
inherently leads to a partial compartmentalization of the ventricular
system, with large-scale vortex structures in the anterior and middle
ventricles~(~\Crefpanels{fig:fig1}{e}). The highest velocities ($26.5$ \textmu m/s)
occur in the vicinity of the dorsal cilia in the middle ventricle.

To examine the effect of cardiac pressure pulsations,
we next consider a no-cilia model variation. In this scenario, the vortex structures
generated by the ciliary motion vanish. The simulated CSF
flow is pulsatile as a result of the sinusoidal pressure gradient (~\Crefpanels{fig:fig1}{f, g}),
with a peak velocity magnitude of $4.8$ \textmu m/s. At each time instant, the flow
is unidirectional, with CSF initially flowing rostrocaudally (positive $x-$axis) before changing
direction to caudorostral in the middle of the cardiac cycle.

Finally, we consider the baseline model, simulating flow driven by both cilia
and cardiac pulsations. We observe persistent flow features,
with vortex-structures in the anterior and middle ventricles
and directional flow in the posterior ventricle (~\Crefpanels{fig:fig1}{h--j},
\fixme{Supplementary video 1}). These characteristic patterns suggest
that the flow in the anterior and middle ventricles is majorly driven
by cilia, while the cardiac pressure pulsations dominate
within the posterior ventricle and in the ducts connecting the ventricular
compartments. Over one cardiac cycle, the CSF flow
speed reaches up to 27.7 \textmu m/s, dorsally in the posterior part
of the middle ventricle. This peak is attained at the time instant
coinciding with the time of peak caudal cardiac pulsatile flow, since 
both the tangential traction induced by the cilia and the pulsatile
pressure-driven flow are aligned with the rostrocaudal 
axis. Using the observed maximum velocity magnitude and choosing the height of
the middle ventricle ($110$ \textmu m) as a characteristic length
scale yields a Reynolds number $\mathrm{Re} \approx 0.004$,
justifying the assumption of Stokes flow. In summary, the computational
CSF flow model reproduces the flow features observed
in vivo~\cite{Olstad2019CiliaryDevelopment}.

\subsection{Ventricular solute transport is advection-dominated for larger molecules}
We next explore the impact of these CSF flow patterns and magnitudes
on solute transport within the brain ventricles.
To this end, we simulate the distribution and evolution
of different solutes subjected to photoconversion within a region of
interest in the dorsal region of the middle ventricle (ROI 1)
(~\Crefpanels{fig:fig2}{a}). We consider the spread of
a solute concentration with a given molecular diffusion coefficient,
corresponding to exosomes ($D_1 = 1.63 \times 10^{-12}$ m$^2$/s),
a circa 90 kDa protein STM-GFP ($D_2 = 5.75 \times 10^{-11}$ m$^2$/s),
and the 26 kDa photoconvertible protein  Dendra2 ($D_3 = 1.15 \times 10^{-10}$ m$^2$/s).
As expected, transport occurs more
rapidly with increasing diffusion coefficient. Interestingly, the
distribution patterns differ between the solutes.
For the smallest diffusion coefficient, representing
exosomes, the advective transport by the CSF flow dominates diffusion
(\Crefpanels{fig:fig2}{b}). The vortex structures of the velocity field
carry over into the concentration field, indicating that transport
along the streamlines is more rapid than the diffusion across the
streamlines. The balance between diffusion and advection shifts with
increasing diffusion coefficients: the CSF flow structures become more
obfuscated, and the solute spreads more uniformly throughout the
ventricular geometry (\Crefpanels{fig:fig2}{c, d}). Notably, the
smaller solutes with higher diffusion coefficients spread more quickly
to regions distal to the photoconversion site.
\begin{figure}%[H]
    \centering
    \includegraphics[width=\textwidth]{graphics/figure2_simulation_results_different_coeffs.png}\caption{Simulated distribution and evolution after photoconversion of
    exosomes ($D_1$), STM-GFP ($D_2$) and Dendra2 ($D_3$). 
    \textbf{a.} Geometry-centered clip of the ventricles mesh showing the locations of the ROIs.
    \textbf{b--d.} Slice of the concentration field simulated
    with diffusion coefficients $D_1$, $D_2$ and $D_3$ for eight time instants covering the
    time range from 45 seconds to 790 seconds, corresponding to 100 to 1750 cardiac cycles.
    Slices are taken in the $xz$-plane ($y=0.140$ mm, the center of the geometry).
    The colorbar is scaled so the maximum is the maximum concentration of the Dendra2 
    concentration field slice at the last time instant.
    \textbf{e--g.} Mean concentration $\cbar$ in each region of interest (ROI)
    as function of time for diffusion coefficient $D_1$, $D_2$ and $D_3$, respectively. 
    \textbf{h.} The time $\hat{t}$ before the mean concentration $\cbar$ exceeds a
    threshold value of $0.25$ (ROIs 2-4) or $0.10$ (ROIs 5, 6) in each simulation setup.
    Note that there is no bar for $D_1$ in ROI 6, because the threshold value was never exceeded.}\label{fig:fig2}
\end{figure}

To quantify the transport of solute, we plot the profile of the mean
concentrations $\cbar_i(t)$ as functions of time
(\Crefpanels{fig:fig2}{e--g}) and
the times $\hat{t}$ when the mean concentrations $\cbar_i$ first exceed
a threshold value $\hat{c}$ (\Crefpanels{fig:fig2}{h})
in 6 ROIs (\Crefpanels{fig:fig2}{a}).
The dynamics at the site of photoconversion (ROI 1)
is logarithmic and follows from imposing \eqref{eq:photoconversion_curve}.
For ROIs 2--6, we observe faster kinetics with higher diffusion coefficients
(\Crefpanels{fig:fig2}{e--g}),
which results in smaller times-to-threshold
(\Crefpanels{fig:fig2}{h}).
For all three diffusion coefficients, the dynamics of ROI 5
is significantly delayed compared to ROI 2--4,
and ROI 6 is the slowest. Notably, for $D_1$ the mean concentration
in ROI 6 is close to zero at the end of simulation.
To quantify the balance of advection versus diffusion as transport mechanisms,
we calculate the Péclet numbers $\mathrm{Pe}_1=519$,
$\mathrm{Pe}_2=14.7$ and $\mathrm{Pe}_3=7.4$ for $D_1$, $D_2$ and $D_3$, respectively.
Since $\mathrm{Pe}>1$, this indicates that advection is the
dominant transport mechanism on a global scale.

\subsection{Validation of the simulations with experimental data}
To validate our computational model, we performed photoconversion experiments
in transgenic zebrafish larvae expressing the secreted Dendra2 in their brain ventricles.
Our photoconversion protocol consisted of the acquisition of a baseline,
before localized exposure to a UV laser (at the ROI 1 location)
that converts Dendra2 from a green to a red emitting fluorescent protein. 
To quantify the transport of photoconverted proteins, we measured and
calculated the change in fluorescence intensity $\Delta F = (F(t)-F_0)/F_0$
over time (\Crefpanels{fig:fig3}{a},
$F_0$ being the average fluorescence at baseline).

We obtained very similar kinetics curves (\Crefpanels{fig:fig3}{a, b})
and time-to-threshold values (\Crefpanels{fig:fig3}{c, d}) when comparing our
experiments (\Crefpanels{fig:fig3}{a, c}) with
simulation (\Crefpanels{fig:fig3}{b, d}).
Importantly, since the initial condition of the simulation is a mean concentration of zero,
the mean concentration $\cbar$ can be interpreted as the change in mean concentration and
is thus a similar metric to $\Delta F$.
The final values of $\cbar$ in our simulations 0.76, 0.72, 0.82, 0.56 and 0.35
(ROIs 2--6, respectively)
are generally higher than the mean experimental values 0.63, 0.71, 0.67, 0.29 and 0.18 for $\Delta F$.
Notwithstanding similar dynamics, we observe small differences between the experiments and simulation.
For instance, transport in ROIs 2--6 is faster in the simulation and reaches higher final values.
Moreover, the $\cbar$ profiles of ROIs 3--4 are opposite order in
the simulations as compared to the experiments.
Despite observing small differences, the in vivo experiments and
simulations are similar enough to convincingly validate the model.
\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{graphics/figure3_compare_experiments_and_simulations.png}
    \caption{
    \textbf{a.} Relative change in fluorescence intensity $\Delta F(t)$ 
    in each region of interest (ROI)
    observed during experiments of photoconversion of Dendra2 proteins in ROI 1.
    \textbf{b.} Mean concentration $\cbar(t)$ in each ROI in a
    simulation of Dendra2 protein transport with
    the diffusion coefficient $D_3$.
    \textbf{c.} Time to $\Delta F(t)$ reaches a threshold value of 0.25 in
    one representative zebrafish larva.
    Time to $\cbar(t)$ reaches a threshold value of 0.25 for
    a transport simulation with diffusion coefficient $D_3$.
    We have excluded the posterior-most regions of the geometries,
    because the values of $\Delta F$ and $\cbar$ never exceeded the threshold
    value of 0.25 in these regions.}
    \label{fig:fig3}
\end{figure}
% Final Delta F values
%0.6322650156762044 ROI 2
%0.7059449124175479 ROI 3
%0.6694500367160006 ROI 4
%0.28986700094045437 ROI 5
%0.18155055624229066 ROI 6
% Final mean c values ROIs 2-6: 0.7637806  0.72025859 0.81903216 0.56204164 0.34926411

\subsection{Absence of ciliary motion affects local protein distribution}
Since CSF flow is highly disturbed in the absence of motile cilia, we
want to assess the impact of cilia motion on solute distribution. To this end,
we measured the movement of photoconverted Dendra2 signal
in control and motile-cilia mutant (\emph{smh}) zebrafish .
By plotting the means of $\Delta F$ (\Crefpanels{fig:fig4}{a})
and the time to threshold
across the control and mutant cohorts (\Crefpanels{fig:fig4}{b}),
we identified a significant
difference in Dendra2 distribution in ROIs 2 and 4, but not in ROIs 3, 5, and 6. In ROIs 2 and 4,
which measure Dendra2 distribution downstream of the photoconversion site,
we observed a slower increase of fluorescence intensity,
resulting in $45\%$ (ROI 2) and $25\%$ (ROI 4)
higher mean times-to-threshold in the \emph{smh}
mutants compared to controls (\Crefpanels{fig:fig4}{b}).

Next, to identify whether our model would recapitulate these findings,
we simulate protein transport by using the only-cardiac/no-cilia flow model 
and compare it with the baseline model
(both cilia and cardiac motion) (\Crefpanels{fig:fig4}{c}).
Similar to our experiments, we observe slower dynamics in simulations for ROIs 2 and 4,
but small changes for ROIs 3 and 6.
Interestingly, the difference was even greater for the simulations,
with the times-to-threshold in ROI 2 and 4 being 90\% and 34\% higher
in the absence of cilia-mediated flow.
Moreover, ROI 5 exhibits very different $\cbar$ profiles for the two
configurations in our computational model,
but not in our experimental data.
Altogether, our findings identify that diffusion plays an important
role for solute movement to regions upstream in the cilia flow fields or to distant ventricles.
\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{graphics/figure4_compare_control_mutant.png}
    \caption{\textbf{a.} Relative change in fluorescence intensity $F(t)$
    in each region of interest (ROI) for control (black lines)
    and \emph{smh} mutant (red lines) fish. 
    \textbf{b.} Times when $\Delta F(t)$ first exceeds a threshold value,
    equal to 0.25 (ROIs 2--4) or 0.10 (ROIs 5, 6), for control (gray dots) and
    \emph{smh} mutant (red dots) fish.
    The thick black lines denote mean values of the datasets. The $p$-values
    were calculated using a Mann-Whitney U test with a confidence level of 95\%, with 
    the null hypothesis that the two distributions of control and mutant data are 
    from the same population.
    \textbf{c.} Mean concentration $\cbar(t)$ in each ROI,
    simulated using the baseline model (black lines) and the cardiac-only/no-cilia model (red lines).
    The horizontal gray lines mark the threshold value $\hat{c}$, and the time
    $\hat{t}$ when the threshold is first exceeded is reported in the plot legends.
    Note that in both panels \textbf{a} and \textbf{c}, the vertical axes have
    different limits in each plot.}
    \label{fig:fig4}
\end{figure}
